I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-15 Paper/Paper_noProof.tex
overall

Compensating pair of paths will make precision of the target state reduce when using overabstraction of values. Here, $weight(x_5) \leq 4$ (the value is reached for $x_1=1$ and $x_2=-1$). However, Box computes an upper bound of $6$, and while DeepPoly recovers some of the dependency in $x_2$, it can only cancel out some of $x_2$ value due to the ReLU over-abstraction (contribution of $\frac{-x_2+4}{2} \leq 2.5$ instead of exact $-x_2 \leq 1.5$), obtaining a bound $weight(x_5) \leq 5.5$ instead of $4$.

@@

We experiment with the smallest network (MLP $6\times 100$), so that we can obtain exact bounds for the first few layers using a full MILP encoding of the DNN. We test over the $\vx=59$th image in the MNIST dataset, as it has a large number of unstable ReLU nodes in the first few layers ($61$ in the first layer and $94(?)$ in the second layer), so we can experiment with a larger choice of values for $K$.

@@

\begin{table}[t!]
	\centering
	\begin{tabular}{|c||c|c|}
	\hline
	\text{Number $K$ of nodes in $Z$}  &  \text{Compensate strength} & \text{Random Choice}  \\ \hline
	\hline
	0  &  2.22 & 2.22  \\ \hline
	10  &  1.84 & 2.03  \\ \hline
	20  &  1.50 & 1.82  \\ \hline
	30  &  1.28 & 1.62  \\ \hline
	40  &  1.14 & 1.44  \\ \hline
	50  &  1.06 & 1.23  \\ \hline
	61 (max) & 1.05 &  1.05 \\ \hline
\end{tabular}
\caption{Average uncertainty of $\MILP_Z$ for nodes of the second layer, for a set $Z$ of $K$ ReLU nodes of the first layer with the heaviest compensation, vs $K$ random nodes of the first layer.}
\label{tab:example0}
\end{table}

@@

We report in Table \ref{tab:example1} the average uncertainty of $\MILP_Z$ following the choice of the $K$ heaviest compensating ReLU nodes in $Z$, vs choosing $K$ nodes randomly from layers $1$ and $2$. 
We compare choosing nodes in the second layer only with choosing nodes in both the first and second layer.
Choosing ReLU nodes in the previous layer (layer 2) only is never a good choice. 
When picking in both layer 1 and 2, choosing 50 out of 149(?) unstable ReLU nodes allows to reduce by $90\%$ the average uncertainty due to inacurate computations ($0.059$ vs $0.574$), while it only reduces the inacuracy by $65\%$ when choosing nodes randomly. Notice that this is direct inacuracy, applying this layer over layer will even further reduce inacuracy created by accumulation of inacuraccies.

@@

\begin{table}[t!]	
	\centering
	\begin{tabular}{|c||c|c|c|}
	\hline
	\text{Number $K$}  &  \text{Compensate layer} 1+2 &  \text{Compensate layer} 2 & \text{Random layer } 1+2 \\ \hline
	\hline
	0  &  1.758 & 1.758 & 1.758  \\ \hline
	10  &  1.677 & 1.677 & 1.695  \\ \hline
	20  &  1.539 & 1.567 & 1.622  \\ \hline
	30  &  1.431 & 1.489 & 1.542  \\ \hline
	40  &  1.329 & 1.454 & 1.47  \\ \hline
	50  &  1.243 & 1.432 & 1.39  \\ \hline
	max  &  1.184 & 1.422 & 1.184  \\ \hline
\end{tabular}
\caption{Average uncertainty of $\MILP_Z$ for nodes of the third layer, for a set $Z$ of $K$ ReLU nodes of the (first and) second layer with the heaviest compensation, vs $K$ random nodes of the first and second layer.}
\label{tab:example1}
\end{table}

@@

	\item We do not report results from kPoly \cite{kpoly}, OptC2V \cite{optC2V}, 
	SDPFO \cite{SDPFI} which do not present favorable trade offs compared with $\beta$-CROWN. Results can be found in \cite{crown} and are performed in the same conditions so can be compared.
	\item Finally, we run several version of Compensate($K$). First, we have a fixed choice for $K$ ($K=60$ for fast and $K=200$ for slow), with a fall back to a lower $K$ ($K=50$ for fast and $K=60$ for slow) if the bounds found is too far from the solution found (MIPGap $>0.5$ in Gurobi). We run (our implementation of) DeepPoly first, then the fast version (fast), before running slow on the remaining Images (fast+slow).	
	Further, we also experiment with an {\em adaptative} implementation, which lowers number $K$ adaptively between layers to adapt to the growing number $N$ of linear variables, which has an impact on the overall $\MILP(N,K)$ runtime. This is particularly useful for deeper network where $N$ is very different between the first and the last layer. Again, we run first an aggressive version of fast, before going to a more concervative slow version to treat the remaining images. All versions compute bounds for 20 neurons of the same layer in parallel, no GPU used.
\end{itemize}

@@

\begin{table}[t!]
	\centering
	\begin{tabular}{|c||c|c|cc||c||c|}

		\hline
		\text{DNN}  & DeepPoly & PRIMA &  \multicolumn{2}{@{}c@{}|}{\text{ $\beta$-CROWN (pureBaB)}} & 
		Comp(adapt.) & Upper\\ 
		 & / CROWN & (refined) & TO=30s & TO=1200s & fast & Bound\\
		\hline \hline

		$6\times500$\  &   $26\%$ & &  &  $44.5\%$  & $\mathbf{65\%}$ & $92\%$ \\

		$\epsilon = 0.035$ & 34s & &  & 576s & 925s &\\ \hline
	\end{tabular}
	\caption{$\%$ of verified images and average runtime in seconds, over 200 images.}
	\label{tab:example3}
	\vspace{-0.8cm}
\end{table}

@@

\begin{figure}[b!]
\centering
\includegraphics[scale=0.8]{Graph.png}
\end{figure}

>>>>

Compensating pair of paths will make precision of the target state reduce when using overabstraction of values. Here, $weight(x_5) \leq 4$ (the value is reached for $x_1=1$ and $x_2=-1$). However, Box computes an upper bound of $6$, and while DeepPoly recovers some of the dependency in $x_2$, it can only cancel out some of $x_2$ value due to the ReLU over-abstraction (contribution of $\frac{-x_2+4}{2} \leq 2.5$ instead of exact $-x_2 \leq 1$), obtaining a bound $weight(x_5) \leq 5.5$ instead of $4$.

@@

We experiment with the smallest network (MLP $6\times 100$), so that we can obtain exact bounds for the first few layers using a full MILP encoding of the DNN. We test over the $\vx=59$th image in the MNIST dataset, as it has a large number of unstable ReLU nodes in the first few layers ($61$ in the first layer and $55$ in the second layer), so we can experiment with a larger choice of values) for $K$.

@@

\begin{table}[t!]
	\centering
	\begin{tabular}{|c||c|c|}
	\hline
	\text{Number $K$ of nodes in $Z$}  &  \text{Compensate strength} & \text{Random Choice}  \\ \hline
	\hline
	0  &  2.22 & 2.22  \\ \hline
	10  &  1.84 & 2.03  \\ \hline
	20  &  1.50 & 1.82  \\ \hline
	30  &  1.28 & 1.62  \\ \hline
	40  &  1.14 & 1.44  \\ \hline
	50  &  1.06 & 1.23  \\ \hline
	61 (max) & 1.05 &  1.05 \\ \hline
\end{tabular}
\caption{Average uncertainty of $\MILP_Z$ for nodes of the second layer, for $Z$ with $K$ ReLU nodes of the first layer (compensating strenght vs random choice).}
\label{tab:example0}
\vspace{-0.8cm}
\end{table}

@@

We report in Table \ref{tab:example1} the average uncertainty of $\MILP_Z$ following the choice of the $K$ heaviest compensating ReLU nodes in $Z$, vs choosing $K$ nodes randomly from layers $1$ and $2$. 
We compare choosing nodes in the second layer only with choosing nodes in both the first and second layer.
Choosing ReLU nodes in the previous layer (layer $2$) only is never a good choice. 
When picking in both layer $1$ and $2$, choosing $50$ out of $116$ unstable ReLU nodes allows to reduce by $90\%$ the average uncertainty due to inacurate computations ($0.059$ vs $0.574$), while it only reduces the inacuracy by $65\%$ when choosing nodes randomly. Notice that this is direct inacuracy, applying this layer after layer will even further reduce inacuracy created by accumulation of inacuraccies.

@@

\begin{table}[t!]	
	\centering
	\begin{tabular}{|c||c|c|c|}
	\hline
	\text{Number $K$}  &  \text{Compensate layer} 1+2 &  \text{Compensate layer} 2 & \text{Random layer } 1+2 \\ \hline
	\hline
	0  &  1.758 & 1.758 & 1.758  \\ \hline
	10  &  1.677 & 1.677 & 1.695  \\ \hline
	20  &  1.539 & 1.567 & 1.622  \\ \hline
	30  &  1.431 & 1.489 & 1.542  \\ \hline
	40  &  1.329 & 1.454 & 1.47  \\ \hline
	50  &  1.243 & 1.432 & 1.39  \\ \hline
	116 &  1.184 & 1.422 & 1.184  \\ \hline
\end{tabular}
\caption{Average uncertainty of $\MILP_Z$ for nodes of the third layer, for $Z$ with $K$ ReLU nodes of the 1st and 2nd layer (compensating strenght vs random choice).}
\label{tab:example1}
\vspace{-0.8cm}
\end{table}

@@

	\item We do not report results from kPoly \cite{kpoly}, OptC2V \cite{optC2V}, 
	SDPFO \cite{SDPFI} which do not present favorable trade offs compared with $\beta$-CROWN. Results can be found in \cite{crown} and are performed in the same conditions so can be compared.
	\item Finally, we run several version of Compensate($K$). First, we have a fixed choice for $K$ ($K=60$ for fast and $K=200$ for slow), with a fall back to a lower $K$ ($K=50$ for fast and $K=60$ for slow) if the bounds found is too far from the solution found (MIPGap $>0.5$ in Gurobi). We run (our implementation of) DeepPoly first, then the fast version, before running slow on the remaining Images.
	Further, we also experiment with an {\em adaptative} implementation, which lowers number $K$ adaptively between layers to adapt to the growing number $N$ of linear variables, which has an impact on the overall $\MILP(N,K)$ runtime. This is particularly useful for deeper network where $N$ is very different between the first and the last layer. Again, we run first an aggressive version of fast, before going to a more concervative slow version to treat the remaining images. All versions compute bounds for 20 neurons of the same layer in parallel, no GPU used.
\end{itemize}

@@

\begin{table}[b!]
	\centering
	\begin{tabular}{|c||c|c|cc||c||c|}

		\hline
		\text{DNN}  & DeepPoly & PRIMA &  \multicolumn{2}{@{}c@{}|}{\text{ $\beta$-CROWN (pureBaB)}} & 
		Comp(adapt.) & Upper\\ 
		 & / CROWN &  & TO=30s & TO=1200s & fast & Bound\\
		\hline \hline

		$6\times500$\  &   $26\%$ & $35\%$ &  &  $44.5\%$  & $\mathbf{65\%}$ & $92\%$ \\

		$\epsilon = 0.035$ & 34s & 121s &  & 576s & 925s &\\ \hline
	\end{tabular}
	\caption{$\%$ of verified images and average runtime in seconds, over 200 images.}
	\label{tab:example3}
	\vspace{-0.8cm}
\end{table}

@@

\begin{figure}[t!]
\centering
\includegraphics[scale=0.8]{Graph.png}
\end{figure}
